---
title: "compatibility"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{compatibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Temporary working directory for example data package

twd <- tempfile('vpkg')
dir.create(twd)
knitr::opts_knit$set(root.dir = twd)
```

### Converting a DPR1 Datapackage to DPR2 
Sometimes users may encounter datapackages created by DPR1 and still want to use DPR2 functionalities. DPR2 has function `dpr_dpr1_yaml_convert` which converts a DPR1 YAML object into a DPR2 compatible yaml object

```{r}
#existing dpr1 project
library(DPR2)
library(DataPackageR)
library(yaml)

#setup an empty DPR1 skeleton
processing_code <-
    system.file("extdata", "tests", "subsetCars.Rmd", package = "DataPackageR")

# Create the package framework.
DataPackageR::datapackage_skeleton(
  name = "mtcars20",
  code_files = processing_code,
  r_object_names = "cars_over_20"
)

# set working directory for this vignette to the newly created datapackage
knitr::opts_knit$set(root.dir = file.path(getwd(), 'mtcars20'))

# create YML
yml <- DataPackageR::construct_yml_config(code = "subsetCars.Rmd", data = "cars_over_20")

cat(yaml::as.yaml(yml))
```



After using dpr_dpr1_yaml_convert function and loading and looking over the yaml settings, we see the yaml configuration for the DPR1 package is converted to DPR2 compatible settings:

```{r}
DPR2::dpr_dpr1_yaml_convert()
dpr_yaml_get()
```
