---
title: "DataPackageR Compatibility"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DataPackageR Compatibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Converting a DataPackageR data package to use DPR2

Sometimes users may encounter datapackages created by DataPackageR. DPR2 has the function `dpr_convert` which converts a DataPackageR package into a DPR2 package. In this example we will copy out DataPackageR test package to a temp directory to demonstrate the DPR2 conversion function.

```{r echo = T}
file.copy(
  "../tests/testthat/dpr1package/",
  tempdir(),
  recursive = TRUE
)
```

To convert a DataPackageR package to use DPR2 simply run `dpr_convert(path)`, setting the path to the DataPackageR package, whereas the default path is the current working directory.

```{r}
## copy test package to temp directory
file.copy(
  "../tests/testthat/dpr1package/",
  tempdir(),
  recursive = TRUE
)

dpr1_path <- file.path(tempdir(), "dpr1package")

## load DPR2 
library(DPR2)

## convert test DataPackageR test package to the DPR2 format 
dpr_convert(dpr1_path)
```

Notice that the above conversion does do a bit of digest validation, where in the above example on of the tracked files has a different checksum than what is reported in the digest. 

Converting a package using `dpr_convert()` deletes several files created by DataPackageR that DPR2 does not use. 

* `DATADIGEST`
* `NEWS.md`
* `Read-and-delete-me`,
* `R/documentation.R`
* `data-raw/documentation.R`,
* `inst/extdata/Logfiles`, 
* any file in `inst/doc` with the same name as an `R` or `Rmd` file found in the `data-raw` or`Logfiles`

`dpr_convert()`  also modifies the package `DESCRIPTION`, removeing its `Date`, and `DataVersion` fields.

Although the `DATADIGEST` file is deleted, DPR2 does maintain that information, but uses a directory found in `inst` to do so, and DPR2 provides the `dpr_data_digest` function for assembling the contents of the new data digest format into a tablular object. See [Data Versioning](data_versioning.html) for more information regarding how to use the new data digest format.

```r
## view the new data digest
dpr_data_digest(dpr1_path)
```

`dpr_convert` also translates the orginal `DataPackageR` yaml file into a DPR2 compatible format. Examining the `datapackager.yml` file, we see the yaml configuration for the DataPackageR package is converted to DPR2 compatible settings.

```r
## view the new yaml file
dpr_yaml_get(dpr1_path)
```

DataPackageR used the `data-raw` directory to store data processing scripts. DPR2 uses a new `processing` directory to store those, however this is configurable in the DPR2 `datapackager.yml` file, and thus converted data packages has this set to the default DataPackageR `data-raw` path.,
