---
title: "Data Versioning"
author: "Jason Taylor"
date: "2025-03-26"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Versioning}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

DPR2 offers features to facilitate tracking data versioning throughout a project's lifecycle. This vignette will cover the those features.

```{r echo = F}

library(DPR2)
dpr2 <- "../tests/testthat/dpr2package"
path <- file.path(tempdir(), "dpr2package")
file.copy(dpr2, tempdir(), recursive = TRUE)

git2r::init(path = path, branch="main")
git2r::add(repo = path, path = ".")
Sys.sleep(1) # odb_blobs appears to be ordered by when, and the smallest unit is 1 sec, adding pauses to get correct sorting
git2r::commit(repo = path, message="init commit")

dpr_build(path)
git2r::add(repo = path, path = ".")
git2r::commit(repo = path, message="Our first build")

irisPath <- file.path(path, "processing/iris.R")
irisScript <- readLines(irisPath)

irisScript[
  which(irisScript == "iris$Sepal.Area <- iris$Sepal.Length *  iris$Sepal.Width")
]<-  "iris$Petal.Area <- iris$Petal.Length * iris$Petal.Width"

writeLines(irisScript, irisPath)
dpr_build(path)

git2r::add(repo = path, path = ".")
Sys.sleep(1)
git2r::commit(repo = path, message="Change sepal area to petal area.")

mtcarsPath <- file.path(path, "processing/mtcars.R")
mtcarsScript <- readLines(mtcarsPath)
mtcarsScript[which(mtcarsScript == "mazda <- mtcars[grep(\"Mazda\", row.names(mtcars)),]")] <- "mazda <- mtcars[grep(\"Mazda\", row.names(mtcars)),][1]"
writeLines(mtcarsScript, mtcarsPath)
dpr_build(path)

git2r::add(repo = path, path = ".")
Sys.sleep(1)
git2r::commit(repo = path, message="Update to mazda cars.")

treesPath <- file.path(path, "processing/trees.Rmd")
treesScript <- readLines(treesPath)
treesScript <- c(
  treesScript[1:which(treesScript == "plot(trees)")],
  "trees <- trees[1:3,]",
  treesScript[(which(treesScript == "plot(trees)")+1):length(treesScript)]
)
writeLines(treesScript, treesPath)
dpr_build(path)

git2r::add(repo = path, path = ".")
Sys.sleep(1)
git2r::commit(repo = path, message="Trees subsetting")

```

Printing the data object history from a data package under version control and comparing those to the current version of the data pacakge.

```{r}

history <- dpr_data_history(path=path, include_checksums=TRUE)
print(history)

firstIrisHash <- history[history$name == "irisArea.rda",
                     ][1,
                       ]$blob_git_sha1
print(firstIrisHash)
firstIris <- dpr_recall_data_versions(firstIrisHash, path)

lastIrisHash <- history[history$name == "irisArea.rda",
                    ][nrow(history[history$name == "irisArea.rda",]),
                      ]$blob_git_sha1
print(lastIrisHash)
lastIris <- dpr_recall_data_versions(lastIrisHash, path)

digest::digest(firstIris[[1]][[1]])
digest::digest(lastIris[[1]][[1]])

dpr_data_digest(path)

```

Rendering a package before building it and comparing the object checksums between the refreshed object in the data directory and the the last recorded checksums in the data digest.

```{r}

irisPath <- file.path(path, "processing/iris.R")
irisScript <- readLines(irisPath)

irisScript <- c(
  irisScript[1:which(irisScript == "iris$Petal.Area <- iris$Petal.Length * iris$Petal.Width")],
  "iris$Sepal.Area <- iris$Sepal.Length * iris$Sepal.Width",
  irisScript[(which(irisScript == "iris$Petal.Area <- iris$Petal.Length * iris$Petal.Width") + 1): length(irisScript)]
)

writeLines(irisScript, irisPath)

```

And once there may have been changes, we can compare datasets.

```{r}

dpr_render(path)

dpr_compare_data_digest(path)

```

